Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures

### 2011-12-30 ltx/paper.tex
re-did performance metrics for pets2000 and pets2001. re-wrote GPUDDPM part of paper.

\documentclass[smallcondensed, final]{svjour3}
% \documentclass[smallcondensed, final, numbook]{svjour3}[12/23/2011]
% \documentclass[twocolumn, final]{svjour3}

@@

\title{Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures}
\titlerunning{Unsupervised Object Detection and Tracking with Dependent Dirichlet Processes}
\authorrunning{Willie Neiswanger and Frank Wood}
% \subtitle{A Bayesian Nonparametric Approach to Tracking Multiple Objects in Videos}
% \subtitle{Subtitle?}
\author{Willie Neiswanger$^{*}$ and Frank Wood$^{\dagger}$}
\institute{
	$^{*}$Columbia University, Department of Applied Math and Applied Physics. 
	Tel.: 503-464-6152.
	\email{wdn2101@columbia.edu}
	\and
	$^{\dagger}$Columbia University, Department of Statistics. 
	Tel.: 212-851-2150. 
	\email{fwood@stat.columbia.edu}}
\date{}

@@

\begin{abstract}
This paper proposes an unsupervised technique for the detection and tracking of arbitrary objects in videos. It is intended to reduce the need for detection or localization methods tailored to specific objects and serve as a general framework applicable to videos with varied object types, backgrounds, and film qualities. The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be extracted in a general manner from the regions in a video that represent objects. This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking. This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track variety of objects with diverse physical charactersitics over non-uniform backgrounds and through occlusion.
% \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
\end{abstract}

\section{Introduction}
% % \subsection{Topic Overview and Terminology}

@@

Research in the past ten years involving the development of methods for general detection and tracking of objects has tended to focus on one of either extraction, localization, or tracking; attempts to integrate all three tasks and produce a system aimed at multiple arbitrary objects and diverse video types was not often a primary focus, or was attempted in an ad-hoc manner to test or apply a more specific technique. Research aimed at developing methods to accomplish the three tasks in a cohesive manner has been pursued more recently to produce algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking---or in short, that can determine the spatiotemporal region of a video occupied by each distinct object. A widely applied, though unsophisticated, example of general detection and tracking is the technique of blob tracking. Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently---for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into ``blobs'' to accomplish localization. A number of tracking techniques based on finding blobs in consecutive frames of a video with similar locations and appearances have been attempted to maintain the identity of detected blobs over time. The main challenges, however, associated with blob tracking involve the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}. These difficulties stem from the typical blob tracker's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking while resisting the collapse of independent tracks.

This paper presents the use of a Bayesian nonparametric mixture model for general, unsupervised multiple object detection and tracking. This method is intended to provide a widely applicable way to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy---and better maintain the isolation of independent objects during tracking. Specifically, this work describes how a Dirichlet process mixture model can accomplish these goals when it is applied to data that may be extracted in a general manner from videos. We will demonstrate how to extract data from objects in videos and how this generative model allows for automated detection and tracking. The model is first presented in an abstract form, to allow it to be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies. We then formulate a specific model for data gained during a basic extraction procedure and describe inference algorithms that allow object localization and tracking to be carried out. We demonstrate our implementation on a number of synthetic and benchmark datasets and describe and compute performance metrics to quantify results.

@@

Extraction is related to the tasks of foregound segmentation, background modeling, motion detection, and appearance feature extraction. In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene. Areas of research in this field include attempts to handle situations where the background displays motion or other dynamic characteristics, reduce noise caused by the false extraction of backgrounds, handle abberant image effects present in some videos, and operate accurately on scenes involving illumination changes. These procedures can usually be classified as either model-based methods or heuristic-based methods. Model-based methods emphasize the detection of shapes of regions containing objects accurately and reducing background noise. For example, foreground extraction has been robustly carried out by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel at that point \citep{stauffer_1999, elgammal_2000, elgammal_2002}. Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids). Heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics. These methods are often less accurate than model-based methods, but also far less computationally expensive. Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold. The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.  Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing two subsequent images in a video. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}. In addition to the two techniques mentioned above, there exist other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on the `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}. In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.

Localization is closely related to the topics of object detection, position-finding, segmentation, and target recognition. These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance. There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time. Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation, parsing, and pixel clustering \willie{must cite here}. These topics usually relate to the automated partitioning of images into meaningful regions. Note that the popular phrase `object detection' is also typically used to refer to procedures that carry out extraction and localization; however, it is often reserved to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.

Tracking is related to the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'. In general, these tasks aim to maintain the localization of the position and/or shape of an object over time. Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object.

Methods for tracking single, non-rigid, objects are numerous. Color, for example, has long been used for tracking. These techniques often operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames. One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs. Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other smooth time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}. Additionally, color has been successfully applied as the feature in a technique known a the mean-shift procedure (or the kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}. This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--optimizes the search for the region in a frame which is most similar to a target region in the previous frame. It involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point. Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors. It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}. The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.

@@

% Extraction

%

@@

\section{Model Overview}
\label{sec:modeloverview}
%
This section provides details on the data to be modeled, the Dirichlet process mixture (DPM) model, and the dependent Dirichlet process mixture (DDPM) model. We make explicit the way in which characteristics of video objects can be represented by elements of these models. We formulate the model in a sufficently abstract manner to allow for implementations tailored for specified objects in future studies.

\subsection{Data}
\label{sec:data}
%
We defined the task of extraction to be the unsupervised determination of which spatiotemporal regions of a video constitute objects, and outlined a number of extraction methods in Section~\ref{sec:relatedwork}. Extraction allows us to generate a collection of data points. Each data point can be thought of as a hypothesis about, or noisy version of, the representation of an object; a collection of hypotheses for a given object is assumed to be distributed in some way around the true object representation. A video of multiple objects is therefore assumed to yield data that contains multiple groups of object hypotheses with similar characteristics (i.e. clusters of similar hypotheses).

Determining the regions of each video frame that contain objects allows us to generate these object hypotheses; namely, we let each hypothesis be a description of the pixel data over an area centered at a point in the extracted region. We can, for example, let each pixel in the extracted region be a hypothesis, and let the features of the representation consist of the spatial position of the pixel. An extraction dataset of this form is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in an extracted region in a given frame, and $t$ is the discrete time index of the frame. These three features are always available once regions containing objects are discerned, and are useful features for modeling (as position information is key for reasonable tracking and time information is necessary for maintaining the temporal dependencies in the data).

Examples of additional features that could potentially be extracted from object-regions in video frames include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform calculations (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and can be represented by distributions that allow for tractable inference (discussed further in Section~\ref{sec:inference}).

We will use the phrase ``baseline'' three-dimensional extraction data to refer to the features $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ of pixels present in regions containing objects. These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when they are plotted in three dimensions (where we let the frame number $t$, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes). An example of data with such a distribution can been seen in Figure~\ref{fig:baselinedata}. The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.

% frank says the following worm figures are a no go
% \begin{figure}[h]
%   \centering
%   \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2001_data_1.png}} 
%   \subfloat[]{\label{fig:baselinedata2}\includegraphics[width=0.33\textwidth]{../img/usbh_data_3.png}} 
%   \subfloat[]{\label{fig:baselinedata3}\includegraphics[width=0.33\textwidth]{../img/traffic2_data_1.png}}\\
%   \subfloat[]{\label{fig:baselinedata4}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_1.png}} 
%   \subfloat[]{\label{fig:baselinedata5}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_2.png}} 
%   \subfloat[]{\label{fig:baselinedata6}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_3.png}}
%   \caption{Baseline extracted features from the (a.) PETS2001 Dataset, (b.) Ants Dataset, (c.) Traffic Dataset, and PETS2009/2010 benchmark dataset, frames (a.) 1-200, (b.) 201-400, (c.) 401-600. Time is shown on the vertical axis. Marker color corresponds with spatial coordinates and is shown for ease of viewing.}
%   \label{fig:baselinedata}
% \end{figure}

@@

\subsection{Dirichlet Process Mixtures}

Dirichlet process mixtures fall under the heading of Bayesian nonparametric (BNP) models. These models have been widely used over the past decade to perform nonparametric density estimation and cluster analysis. They are, in particular, useful for estimating the number of latent classes (clusters) in mixture models. In this work, estimating the number of clusters in the data generated via extraction is shown in the following sections to be equivalent to estimating the number of distinct objects in a video. Consequentially, a model that is able to perform robust nonparametric estimation of the clusters in an extraction dataset may be applied to perform unsupervised object detection and tracking.

The following sections are intended to provide a solid introduction to DPMs without a major foray into the theoretical aspects of the model that require a more rigorus technical setup. We give background on finite mixture models, Bayesian finite mixture models, and Dirichlet processes, which allows us to subsequently describe DPMs in context.

\subsubsection{Finite Mixture Model}

A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (also called `clusters'), where the coefficients of the linear combination sum to one; each mixture component, in turn, is a probability distribution for $x_{i}$, given some parametric form. The finite mixture model can be written as
\begin{equation}
P(x) = \sum_{k=1}^{K} P(c_{i} = k)P(x|\theta_{k})
\end{equation}
where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture (which allows the coefficients of the linear combination to be written as $P(c_{i} = k)$), and $\theta_{k}$ denotes the parametric form of the $k^{\text{th}}$ mixture component. We also define $p_{k} = P(c_{i} = k)$ for $k \in \{ 1, \ldots, K \} $. We can therefore write this model generatively as
\begin{align}
\begin{split}
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}})
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, and the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$).

\subsubsection{Bayesian (Finite) Mixture Model}

The finite mixture model of the previous section can be extended to a Bayesian mixture model by viewing the distribution parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution. In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the K mixture component assignment weights. The resulting Bayesian mixture model can be formulated in a generative representation as
\begin{align}
\begin{split}
\label{bayesian_mixture_model}
	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)\\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
	\theta_{c_{i}}  &\sim  \mathbb{G}_{0} \\
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}})
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$), the $\theta_{c_{i}}$ are drawn from the prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from a Dirichlet prior parameterized by $\alpha/K, \ldots, \alpha/K$.

\subsubsection{Dirichlet Process}

@@

\subsubsection{Dirichlet Process (Infinite) Mixture Model}

>>>>

\documentclass[smallcondensed, final]{svjour3}
% \documentclass[smallcondensed, final]{svjour3}[1/1/2012]

@@

\title{Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures}
\titlerunning{Unsupervised Object Detection and Tracking with Dependent Dirichlet Processes}
% \subtitle{A Bayesian Nonparametric Approach to Tracking Multiple Objects in Videos}

\author{Willie Neiswanger$^{*}$ and Frank Wood$^{\dagger}$}
\authorrunning{Willie Neiswanger and Frank Wood}

\institute{$^{*}$Columbia University, Department of Applied Math and Applied Physics. 
Tel.: 503-464-6152.
\email{wdn2101@columbia.edu}
\and
$^{\dagger}$Columbia University, Department of Statistics. 
Tel.: 212-851-2150. 
\email{fwood@stat.columbia.edu}}

\date{}

@@

\begin{abstract}
This paper proposes a technique for the unsupervised detection and tracking of arbitrary objects in videos. It is intended to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and film qualities. The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be extracted in a general manner from the regions in a video that represent objects. This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking. This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
% \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
\end{abstract}

\section{Introduction}

@@

Research in the past ten years involving the development of methods for general detection and tracking of objects has tended to focus on one of either extraction, localization, or tracking; attempts to integrate all three tasks and produce a system aimed at multiple arbitrary objects and diverse video types was not often a primary focus, or was attempted in an ad-hoc manner to test or apply a more specific technique. Research aimed at developing methods to accomplish the three tasks in a cohesive manner has been pursued more recently to produce algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking---or in short, that can determine the spatiotemporal region of a video occupied by each distinct object. A widely applied, though unsophisticated, example of general detection and tracking is the technique of blob tracking. Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently---for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into ``blobs'' to accomplish localization. A number of tracking techniques based on finding blobs in consecutive frames of a video with similar locations and appearances have been attempted to maintain the identity of detected blobs over time. The main challenges, however, associated with blob tracking involve the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}. These difficulties stem from the typical blob tracker's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking of multiple objects while resisting the collapse of independent tracks.

This paper presents the use of a Bayesian nonparametric mixture model for general, unsupervised multiple object detection and tracking. This method is intended to provide a widely applicable way to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy---and better maintain the isolation of independent objects during tracking. Specifically, this work describes how a Dirichlet process mixture model can accomplish these goals when it is applied to data that may be extracted in a general manner from videos. We will demonstrate how to perform this extraction and how this generative model allows for automated detection and tracking. The model is first presented in an abstract form, to allow it to be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies. We then formulate a specific model for data gained during a basic extraction procedure and describe inference algorithms that allow object localization and tracking to be carried out. We demonstrate our implementation on multiple synthetic and benchmark datasets and compute a few standard performance metrics for each to quantify results.

@@

Extraction is related to the tasks of foregound segmentation, background modeling, motion detection, and appearance feature extraction. In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene. Areas of research in this field include attempts to handle situations where the background displays motion or other dynamic characteristics, reduce noise caused by the false extraction of backgrounds, handle abberant image effects present in some videos, and operate accurately on scenes involving illumination changes. These procedures can usually be classified as either model-based methods or heuristic-based methods. Model-based methods emphasize the detection of shapes of regions containing objects accurately and reducing background noise. For example, foreground extraction has been robustly carried out by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel at that point \citep{stauffer_1999, elgammal_2000, elgammal_2002}. Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids). Heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics. These methods are often less accurate than model-based methods, but also less computationally expensive. Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold. The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.  Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing pairs of consecutive images in a video. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}. In addition, there exist other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on the `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}. In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.

Localization is related to the topics of object detection, position-finding, segmentation, and target recognition. These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance. There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time. Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation and pixel clustering \willie{must cite here}. These topics usually relate to the automated partitioning of images into meaningful regions. Note that the popular phrase `object detection' is also typically used to refer to procedures that carry out extraction and localization; however, it is often reserved to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.

Tracking is related to the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'. In general, these tasks aim to maintain the localization of the position and/or shape of an object over time. Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object. Methods for tracking single, non-rigid, objects are numerous. Color, for example, has long been used for tracking; these techniques often operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames. One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs. Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}. Additionally, color has been successfully applied as the feature in a technique known a the mean-shift procedure (or the kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}. This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--optimizes the search for the region in a frame which is most similar to a target region in the previous frame. It involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point. Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors. It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which allows for clustering \citep{cheng_1995}. The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.

@@

% Extraction
%

@@

\section{Model Overview}
\label{sec:modeloverview}

This section provides details on the extracted data, Dirichlet process mixture (DPM) models, dependent Dirichlet process mixture (DDPM) models, and the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) model. We make explicit the way in which characteristics of video objects can be represented by elements of these models. We formulate the model in a sufficently abstract manner to allow for implementations using specific object representations and extraction procedures in future studies.

\subsection{Data Extraction}
\label{sec:data}

We defined the task of extraction to be the unsupervised determination of which spatiotemporal regions of a video constitute objects, and outlined a number of extraction methods in Section~\ref{sec:relatedwork}. Extraction allows us to generate a collection of data points. Each data point can be thought of as a hypothesis about, or noisy version of, the representation of an object; a collection of hypotheses for a given object is assumed to be distributed in some way around the true object representation. A video of multiple objects is therefore assumed to yield data that contains multiple groups of object hypotheses with similar characteristics (i.e. clusters of similar hypotheses).

Determining the regions of each video frame that contain objects allows us to generate these object hypotheses; namely, we let each hypothesis be a description of the pixel data over an area centered at a point in the extracted region. We can, for example, let each pixel in the extracted region be a hypothesis, and let the features of the representation consist of the spatial position of the pixel. An extraction dataset of this form is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in an extracted region in a given frame, and $t$ is the discrete time index of the frame. These three features are always available once regions containing objects are discerned, and are useful features for modeling (as position and time information is key for reasonable tracking).

Examples of additional features that could be extracted from object-regions in video frames include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform calculations (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and can be represented by distributions that allow for tractable inference (discussed further in Section~\ref{sec:inference}).

We will use the phrase ``baseline'' three-dimensional extraction data to refer to the features $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ of pixels present in regions containing objects. These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when the data are plotted in three dimensions (where we let the frame number $t$, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes). %An example of data with such a distribution can been seen in Figure~\ref{fig:baselinedata}...The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.

% frank says the following worm figures are a no go.
% \begin{figure}[h]
%   \centering
%   \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2001_data_1.png}} 
%   \subfloat[]{\label{fig:baselinedata2}\includegraphics[width=0.33\textwidth]{../img/usbh_data_3.png}} 
%   \subfloat[]{\label{fig:baselinedata3}\includegraphics[width=0.33\textwidth]{../img/traffic2_data_1.png}}\\
%   \subfloat[]{\label{fig:baselinedata4}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_1.png}} 
%   \subfloat[]{\label{fig:baselinedata5}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_2.png}} 
%   \subfloat[]{\label{fig:baselinedata6}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_3.png}}
%   \caption{Baseline extracted features from the (a.) PETS2001 Dataset, (b.) Ants Dataset, (c.) Traffic Dataset, and PETS2009/2010 benchmark dataset, frames (a.) 1-200, (b.) 201-400, (c.) 401-600. Time is shown on the vertical axis. Marker color corresponds with spatial coordinates and is shown for ease of viewing.}
%   \label{fig:baselinedata}
% \end{figure}

@@

\subsection{Dirichlet Process Mixtures}
\label{sec:dpm}

Dirichlet process mixtures fall under the heading of Bayesian nonparametric (BNP) models. These models have been widely used over the past decade to perform nonparametric density estimation and cluster analysis. They are, in particular, useful for estimating the number of latent classes (clusters) in mixture models. In this work, estimating the number of clusters in the data generated via extraction is equivalent to estimating the number of distinct objects in a video. Consequentially, a model that is able to perform robust nonparametric estimation of the clusters in data gained during extraction may be applied to perform unsupervised object detection and tracking.

The following sections are intended to provide an introduction to DPMs without a major foray into the theoretical aspects of the model that require a more rigorus technical setup. We provide background on finite mixture models, Bayesian finite mixture models, and Dirichlet processes, which allows us to subsequently describe DPMs in context.

\subsubsection{Finite Mixture Model}
\label{sec:finite_mixture}

A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (which we also refer to as `clusters'), where each mixture component is a probability distribution for $x_{i}$ with some parametric form, and the coefficients of the linear combination sum to one. The finite mixture model can be written as
\begin{equation}
P(x_{i}) = \sum_{k=1}^{K} P(c_{i} = k)P(x_{i}|\theta_{k})
\end{equation}
$\forall i \in \{ 1, \ldots, N \}$, where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture and $\theta_{k}$ denotes the parameters of the $k^{\text{th}}$ mixture component. Note that by choosing $P(c_{i} = k)$ as coefficients of the linear combination, it is ensured that these coefficients sum to one. We also define $p_{k} := P(c_{i} = k)$ for $k \in \{ 1, \ldots, K \} $. We can therefore write this model generatively as
\begin{align}
\begin{split}
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}})
\end{split}
\end{align}
$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$), and the ``Discrete'' distribution refers to a multinomial distribution whose parameters are a 1-of-K vector (i.e. a vector of counts that sums to one).

\subsubsection{Bayesian (Finite) Mixture Model}
\label{sec:bayesian_mixture}

The finite mixture model of the previous section can be extended to a Bayesian mixture model by viewing parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution. In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the mixture component assignment weights. The resulting Bayesian mixture model can be formulated generatively as
\begin{align}
\begin{split}
\label{bayesian_mixture_model}
	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)\\
		\theta_{k}  &\sim  \mathbb{G}_{0}, \hspace{2mm} \forall k \in \{1, \ldots, K\} \\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}})
\end{split}
\end{align}
$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{k}$ are parameters defining the $k^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{k})$), the $\theta_{k}$ are drawn from a prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from a Dirichlet prior parameterized by $\alpha/K, \ldots, \alpha/K$.

\subsubsection{Dirichlet Process}
\label{sec:dirichlet_process}

@@

\subsubsection{Dirichlet Process (Infinite) Mixture Model}
\label{sec:dirichlet_process_mixture}
